---
title: "Rick and Morty's IMDB Scoring"
author: "Mauro"
date: "2023-12-21"
output: 
  html_document:
    keep_md: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

# starting libraries
library(dplyr)
library(stringr)
library(ggplot2)
library(viridis)
library(lubridate)
library(ggrepel)
```

# Seasons

```{r season, echo=FALSE, message=FALSE, warning=FALSE}
df <- rio::import('rick_morty_imdb.xlsx')

df <- df %>%
    mutate(across('name', str_replace, "\"", "")) %>%
    mutate(across('name', str_replace, '"', "")) %>% 
    mutate(episode_num = row_number()) %>% 
    mutate(cod_episode = paste0('s',season,'e',
                                ifelse(episode < 10, paste0('0', episode), 
                                       as.character(episode)))) %>% 
    mutate(air_date = mdy(air_date)) %>% 
    mutate(season = factor(season))

knitr::kable(
df %>%
    group_by(season) %>%
    summarize(
        mean_imdb = round(mean(rating, na.rm = TRUE),2),
        sd = round(sd(rating, na.rm = TRUE),2)
        ) %>% 
    arrange(desc(mean_imdb))
, "pipe")
```

Using the mean IMDB score of every episode we conclude that best season is #3 and the lowest ranked season is the seventh. This season is also the most "volatile", with the highest standard deviation. 

```{r season_boxplot}
ggplot(df, aes(x=season, y=rating, fill=season)) + 
    geom_boxplot(alpha=0.9) +
    theme(legend.position="none") +
    scale_fill_viridis(discrete = TRUE) + 
    theme_bw()
```

# Episodes

## Top 10 episodes

```{r episodes, echo = FALSE}
knitr::kable(
df %>% 
    slice_max(rating, n = 10) %>% 
    select(name, season, rating)
, "pipe")
```

## Worst 10 episodes

```{r episodes_l, echo = FALSE}
knitr::kable(
df %>% 
    slice_min(rating, n = 10) %>% 
    select(name, season, rating)
, "pipe")
```

# Best director

```{r director}
knitr::kable(
df %>% 
    group_by(directed_by) %>% 
    summarize(
        mean_imdb = round(mean(rating, na.rm = TRUE),2),
        count = n(),
        sd = round(sd(rating, na.rm = TRUE),2)
        ) %>% 
    arrange(desc(mean_imdb))
, "pipe")
```

# Best writers

```{r writers,  warning=FALSE}
dfw <- df %>% 
    select(cod_episode, written_by) %>% 
    mutate(written_by = str_replace(written_by, "&", ",")) %>% 
    mutate(written_by = str_replace(written_by, " ,", ",")) %>%
    tidyr::separate(written_by, c('1', '2', '3', '4', '5', '6'), sep = ',') %>% 
    mutate_if(is.character, str_trim) %>% 
    tidyr::pivot_longer(cols = c('1', '2', '3', '4', '5', '6'), 
                        names_to = NULL,
                        values_to = 'writer') %>% 
    na.omit(writer) %>% 
    left_join(df, by = 'cod_episode')

knitr::kable(
dfw_sum <- dfw %>% 
    group_by(writer) %>%
    summarise(mean_imdb = round(mean(rating),2),
              count = n()) %>% 
    arrange(desc(mean_imdb))
, "pipe")
```

We observe that some of the top writers did only collaborate in one episode, so we'll filter the data in order to check writers of more than two episodes.

```{r writers2}
knitr::kable(
dfw_sum %>% 
    filter(count > 1)
, "pipe")
```

## Rick and Morty IMDB episode scoring over time

```{r graph_over_time, message=FALSE, warning=FALSE}
df %>%
    ggplot( aes(x=episode_num, y=rating, color = season)) +
    geom_line() +
    geom_point() + 
    scale_color_viridis(discrete = TRUE) + 
    theme_bw() + 
    geom_text_repel(aes(label=ifelse(rating>9.39,as.character(name),'')),
                    hjust=0,vjust=1) + 
    geom_text_repel(aes(label=ifelse(rating<6,as.character(name),'')),
                    hjust=0,vjust=-1)

```



