---
title: "Rick and Morty's IMDB Scoring"
author: "Mauro"
date: "2023-12-21"
output: 
  html_document:
    keep_md: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

# starting libraries
library(dplyr)
library(stringr)
library(ggplot2)
library(viridis)
library(lubridate)
library(ggrepel)
```

# Seasons

```{r season, echo=FALSE, message=FALSE, warning=FALSE}
df <- rio::import('rick_morty_imdb.xlsx')

df <- df %>%
    mutate(across('name', str_replace, "\"", "")) %>%
    mutate(across('name', str_replace, '"', "")) %>% 
    mutate(episode_num = row_number()) %>% 
    mutate(cod_episode = paste0('s',season,'e',
                                ifelse(episode < 10, paste0('0', episode), 
                                       as.character(episode)))) %>% 
    mutate(air_date = mdy(air_date)) %>% 
    mutate(season = factor(season))

knitr::kable(
df %>%
    group_by(season) %>%
    summarize(
        mean_imdb = round(mean(rating, na.rm = TRUE),2),
        median_imdb = round(median(rating, na.rm = TRUE),2),
        sd = round(sd(rating, na.rm = TRUE),2)
        ) %>% 
    arrange(desc(mean_imdb))
, "pipe")
```

Using the mean IMDB score of every episode we conclude that best season is #3 and the lowest ranked season is the seventh. This season is also the most "volatile", with the highest standard deviation. 

```{r season_boxplot}
ggplot(df, aes(x=season, y=rating, fill=season)) + 
    geom_boxplot(alpha = .9) +
    geom_dotplot(binaxis = 'y', stackdir = 'center', dotsize = .5, alpha = .75, binwidth = .2) +
    scale_fill_viridis(discrete = TRUE)
```

# Episodes

## Top 10 episodes

```{r episodes, echo = FALSE}
knitr::kable(
df %>% 
    slice_max(rating, n = 10) %>% 
    select(name, season, rating)
, "pipe")
```

## Worst 10 episodes

```{r episodes_l, echo = FALSE}
knitr::kable(
df %>% 
    slice_min(rating, n = 10) %>% 
    select(name, season, rating)
, "pipe")
```

# Best director

```{r director}
knitr::kable(
df %>% 
    group_by(directed_by) %>% 
    summarize(
        mean_imdb = round(mean(rating, na.rm = TRUE),2),
        count = n(),
        sd = round(sd(rating, na.rm = TRUE),2)
        ) %>% 
    arrange(desc(mean_imdb))
, "pipe")
```

# Best writers

```{r writers,  warning=FALSE}
dfw <- df %>% 
    select(cod_episode, written_by) %>% 
    mutate(written_by = str_replace(written_by, "&", ",")) %>% 
    mutate(written_by = str_replace(written_by, " ,", ",")) %>%
    tidyr::separate(written_by, c('1', '2', '3', '4', '5', '6'), sep = ',') %>% 
    mutate_if(is.character, str_trim) %>% 
    tidyr::pivot_longer(cols = c('1', '2', '3', '4', '5', '6'), 
                        names_to = NULL,
                        values_to = 'writer') %>% 
    na.omit(writer) %>% 
    left_join(df, by = 'cod_episode')

knitr::kable(
dfw_sum <- dfw %>% 
    group_by(writer) %>%
    summarise(mean_imdb = round(mean(rating),2),
              count = n()) %>% 
    arrange(desc(mean_imdb))
, "pipe")
```

We observe that some of the top writers did only collaborate in one episode, so we'll filter the data in order to check writers of more than two episodes.

```{r writers2}
knitr::kable(
dfw_sum %>% 
    filter(count > 1)
, "pipe")
```

## Rick and Morty IMDB episode scoring over time

```{r graph_over_time, message=FALSE, warning=FALSE}
df %>%
  ggplot(aes(x = episode_num, y = rating, color = season)) +
  geom_line(size = 1) +  # fixed size, outside aes()
  geom_point(size = 3, shape = 21, stroke = 1, aes(fill = season), color = "black") +
  scale_color_viridis(discrete = TRUE, option = "viridis") +
  scale_fill_viridis(discrete = TRUE, option = "viridis") +
  geom_text_repel(
    aes(label = ifelse(rating > 9.39 | rating < 6, as.character(name), '')),
    color = "white", size = 3, box.padding = 0.35
  ) +
  theme_dark(base_size = 14) +
  theme(
    plot.background = element_rect(fill = "#1e1e1e", color = NA),
    panel.background = element_rect(fill = "#1e1e1e", color = NA),
    legend.background = element_rect(fill = "#1e1e1e", color = NA),
    legend.key = element_rect(fill = "#1e1e1e", color = NA),
    plot.title = element_text(color = "white", size = 16, face = "bold"),
    axis.title = element_text(color = "white"),
    axis.text = element_text(color = "white"),
    legend.text = element_text(color = "white"),
    legend.title = element_text(color = "white")
  ) +
  labs(
    x = "Episode Number",
    y = "Rating",
    color = "Season",
    fill = "Season",
    title = "Ratings over Episodes by Season"
  )
```



